Mixed language processing increases cross-language phonetic transfer in Bengali-English Bilinguals

### 2021-06-08 draft_glossa-psych/glossa-psych_AM-ID_manuscript.tex
edits misc

Bi/multilingual speakers can distinguish between the phonetic norms of their languages \cite{caramazza1973acquisition,macleod2010impact,bosch2003simultaneous} and therefore maintain separate sound categories for each language. However, these categories are not autonomous, and show cross-language influence in both perception and production \cite{flege1995second,fowler2008cross,flege2002assessing}.
%citation
In production, this kind of influence (variously termed transfer, drift, accommodation, interference-- used interchangeably here to indicate any interaction between two sets of phonetic norms) has been studied in two conditions:
(i) While a bilingual speaker is operating in any one of their languages. These studies often compare the speech of bilinguals to monolingual norms, and view transfer as changes to long-term memory representations as a result of acquiring an L2 \citep{guion2003vowel,caramazza1973acquisition,flege1987production}; (ii) When both languages of a bilingual speaker are co-activated. These compare productions during mixed-language processing to participants' own productions while using a single language. The resulting transfer is variously thought to involve short-term memory, online processing costs, language mode, and context-awareness. The present study is concerned with the latter.

Existing research on the phonetic effects of mixed-language production has largely focused on a few pairs of phonologically related languages, and used temporal properties of consonants (in particular, voice onset time or VOT) to measure transfer. However, great variation in results (elaborated in sec.\ref{asymmetries} ) suggests that data from a wider variety of
%elaborate on 'great variation'-- differences between language pairs, plus sociolinguistic situation (effects of language experience etc)-- clearer. 'Types of features'--separate line?- Brief mention - discursive factors in VOT. 
populations, language pairs,  and phonetic features is required in order to make meaningful generalizations. There is no work yet on short-term phonetic transfer in the Indian subcontinent, or in any Indo-Aryan language. Nevertheless, widespread multilingualism in this part of the world suggests that phonetic
%line--rephrase
behavior in these populations can be particularly valuable to understanding the nature of such cross-language interactions, as they are more likely to reflect real-world experience with mixed-language processing.

The present study examines phonetic transfer between Bengali and English in a group of highly proficient bilingual speakers in India. We measure spectral properties (F1 and F2) of two English vowels to ascertain if L1 influence on L2 increases during mixed-language use, relative to a participant's unilingual baseline production of L2. Mixed-language data is elicited in two switching paradigms: code-switching and cued picture-naming. The results demonstrate a dynamic effect of transfer on L2 vowels, but no difference in the extent of transfer between the two paradigms. Findings are discussed in light of recent proposals about asymmetries in transfer effects and the role of connected speech in introducing discursive factors to transfer studies.\\

The rest of this section is structured as follows: \ref{bengali_english_in_india} provides background information on Bengali and Indian English focusing on vowel systems, and suggests on the basis of socio-linguistic facts that for the present population, short-term phonetic interaction during mixed-language use is an ecologically valid paradigm that is likely to be a part of everyday language-use experience. \ref{causes} discusses two potential sources of transfer during language processing-- global co-activation of multiple languages vs the local act of switching between languages. Given the highly multilingual setting that characterizes this population, we note that the former is expected to be a constant feature and therefore any observed phonetic interaction is better understood as resulting from the latter. The design of the present study is motivated on the basis of this discussion. A logical consequence of this model is that such phonetic interaction must be highly localized. \ref{duration} elaborates on this by reviewing literature on the duration and `reversiblity' of transfer effects. A recurring pattern that emerges from existing research on short-term phonetic transfer is that the phonetic influence does not affect both languages of the bilingual speaker equally, although the nature of these reported differences varies greatly across studies. In \ref{asymmetries}, we review some of these findings, focusing on proposed sources of asymmetries, in order to highlight the factors that seem to mediate such phonetic interaction. We argue that these results point to an urgent need for expanding the range of language pairs and sound categories examined in order to make meaningful generalizations about the nature of short-term phonetic transfer. Following from this, \ref{asymmetry between sounds} discusses reported asymmetries between different sound categories of a single language, and develops the hypotheses of the present study on the basis of these. In \ref{paradigms}, we motivate a secondary aim of this study: to verify the postulated differences between two paradigms for eliciting short-term phonetic interaction in an experimental setup, namely code-switching and cued picture-naming. \ref{questions_and_hypotheses} summarizes the research questions and hypotheses.

@@

Indian English (IE) refers to the variety of English that has developed in the Indian subcontinent. In India, it is spoken as an L2 by 129 million people (Census of India, 2011).  English is one of the two official languages, used in education, law, media, as a \emph{lingua franca} mainly for an educated elite in most metropolitan regions, and carries a high prestige value \citep{tollefson2014language}.
%citations 
In spite of the different L1s of its speakers, recent literature suggests that IE has a target phonology that is distinct from any of these, as well as from other native varieties of English (see \cite{sirsa2013effects} for a review). Thus, regional variations are seen as the result of L1-influence on a common underlying target. Excepting a small minority of specific communities \citep{pandey201517}, there are very few L1 speakers of IE.\\

Considering these facts, the present study focuses on short-term phonetic interaction during mixed-language use, because:
\begin{enumerate}[label=(\roman*)]
	\item Since the population is multilingual, we expect long-term representations to be affected by multiple languages. 
	\item Given that examples of an IE phonology without L1 `influence' are very rare, it is more meaningful to think of cross-language transfer in L2 as relative to a speaker's own production in a given `baseline' condition.
\end{enumerate}

@@

The vowel system of IE contains the monopthongs \nt{I, i, E, e, \ae, @/2, a:, O, o, U, u}, represented by the lexical set KIT, FLEECE, DRESS, FACE, and TRAP, STRUT, PALM, LOT/CLOTH, GOAT, FOOT, GOOSE \citep{wells1982accents, masica1972sound}. %Their distribution in the F1XF2 space is shown in (insert table 1.7, Pandey). 
There is a single mid-central vowel corresponding to the categories \nt{2,@,3:}, which are treated as distinct in many native varieties of English. Since the English items used in this study are traditionally transcribed with \nt{2}, we use this symbol to indicate the mid-central vowel throughout. %In IE, the dipthong \nt{@U} is realized as \nt{o:}.

@@

Existing research has distinguished changes in category \textsc{representations} due to the acquisition of multiple sound systems (cf Speech Learning Model \citep{flege1995second,flege2007language}; Perceptual Assimilation Model-L2 \citep{best2007nonnative}), from interaction during processing (transfer vs interference \citep{grosjean2012attempt}; competence vs performance interference \citep{paradis1993linguistic}). 
Given their transient nature (cf sec.\ref{duration}), dynamic changes in production during mixed-language use are generally thought to involve the latter \citep{elias2017effects,simonet2014phonetic}. What triggers this interaction? \cite{olson2016role} argues that while cross-language phonetic effects are largely measured at the point of language switch, it could have two potential sources:
\begin{enumerate}
	\item The local point of switch itself
	\item Global co-activation of two languages -- bilingual language mode \citep{grosjean1998studying} 
\end{enumerate}

@@

\cite{grosjean1998studying} suggests that various aspects of the communicative setting, including exposure to (spoken or written) stimuli in multiple languages and awareness of the interlocutor's being bilingual, could trigger a bilingual mode. Since the participants in this study are in an environment which largely contains mixed-language input, multilingual interlocutors, and no stable ambient language, we expect that phonetic transfer during daily language use, if any, takes place in a bilingual mode.  Thus to preserve ecological validity, we elicited both unilingual and switched utterances in a bilingual language mode. Any observed differences in this paradigm would arguably result from interaction during online processing.

>>>>

Bi/multilingual speakers can distinguish between the phonetic norms of their languages \cite{caramazza1973acquisition,macleod2010impact,bosch2003simultaneous} and therefore maintain separate sound categories for each language. However, these categories are not autonomous-- they show cross-language influence in both perception and production \cite{flege1995second,fowler2008cross,flege2002assessing}. What causes, affects, and results from such influence? Answers to these questions provide important insights into how language `systems' are represented and processed in the mind. %Given that a majority of the world's population is multilingual, this is a central concern for any realistic theory of language.
%show cross-language-- rephrase

%keep just the types here, and put the next 2 lines in the causes&effects section?
Cross-language influence at the level of sounds can be studied broadly in two kinds of conditions:
(i) While a bilingual speaker is operating in any one of their languages, by comparing bilingual speech to monolingual norms \citep[e.g.][]{guion2003vowel,caramazza1973acquisition,flege1987production}; %These studies often compare the speech of bilinguals to monolingual norms, and view the influence as changes to long-term memory representations as a result of acquiring an L2 \citep{guion2003vowel,caramazza1973acquisition,flege1987production}; 
(ii) When both languages of a bilingual speaker are co-activated, by comparing bilingual speakers' mixed-language speech to their own norms while using a single language \citep[e.g.][]{grosjean1994going, bullock2009trying,elias2017effects, simonet2014phonetic}. %These compare productions during mixed-language use to participants' own productions while using a single language. The resulting influence is variously thought to involve short-term memory, online processing costs, language mode, and context-awareness. 
The present study is concerned with the latter. In production, researchers have variously referred to this kind of influence as transfer, drift, accommodation, and interference-- these terms are used interchangeably here to indicate any interaction between two sets of phonetic norms.

Existing research on the phonetic effects of mixed-language production has largely focused on a few pairs of phonologically related languages. A majority of these studies use temporal properties of consonants (in particular, voice onset time or VOT) to measure transfer. However, the reported results vary greatly across studies, and appear to be contingent upon both language-specific features and language experience of the participants. This suggests that data from a wider variety of
%elaborate on 'great variation'-- differences between language pairs, plus sociolinguistic situation (effects of language experience etc)-- clearer. 'Types of features'--separate line?- Brief mention - discursive factors in VOT. 
populations, language pairs,  and phonetic features needs to be considered in order to make meaningful generalizations. There is no work yet on short-term phonetic transfer in the Indian subcontinent, or in any Indo-Aryan language. Nevertheless, widespread multilingualism in this part of the world suggests that phonetic
%line--rephrase
behavior in these populations can be particularly valuable towards understanding the nature of such cross-language interactions, as they are likely to reflect real-world experience with mixed-language processing.\\

The present study examines phonetic transfer between Bengali and English in a group of highly proficient bilingual speakers in India. We measure spectral properties (F1 and F2) of two English vowels to ascertain if L1 influence on L2 increases during mixed-language use, relative to a participant's unilingual baseline production of L2. Mixed-language data is elicited in two switching paradigms: code-switching and cued picture-naming. We compare these results to assess if differences between the paradigms independently influence the outcome of phonetic interaction. The results demonstrate a dynamic effect of transfer on L2 vowels during mixed-language production, but no difference in the extent of transfer between the two paradigms. Findings are discussed in light of recent proposals about asymmetries in short-term phonetic interaction and the role of connected speech in introducing discursive factors to transfer studies.\\

The rest of this section is structured as follows: \ref{bengali_english_in_india} provides background information on Bengali and Indian English focusing on vowel systems. \ref{causes} discusses two potential sources of transfer during language processing-- global co-activation of multiple languages vs the local act of switching between languages. Given the highly multilingual setting that characterizes this population, we note that the former is expected to be a constant feature and therefore any observed phonetic interaction is better understood as resulting from the latter. The design of the present study is motivated on the basis of this discussion. A logical consequence of this model is that such phonetic interaction must be highly localized. \ref{duration} elaborates on this by reviewing literature on the duration and `reversiblity' of transfer effects. A recurring pattern that emerges from existing research on short-term phonetic transfer is that the phonetic influence does not affect both languages of the bilingual speaker equally, although the nature of these reported differences varies greatly across studies. In \ref{asymmetries}, we review some of these findings, focusing on proposed sources of asymmetries, in order to highlight the factors that seem to mediate such phonetic interaction. We argue that these results point to an urgent need for expanding the range of language pairs and sound categories examined in order to make meaningful generalizations about the nature of short-term phonetic transfer. Following from this, \ref{asymmetry between sounds} discusses reported asymmetries between different sound categories of a single language, and develops the hypotheses of the present study on the basis of these. In \ref{paradigms}, we motivate a secondary aim of this study: to verify the postulated differences between two paradigms for eliciting short-term phonetic interaction in an experimental setup, namely code-switching and cued picture-naming. \ref{questions_and_hypotheses} summarizes the research questions and hypotheses.

@@

Indian English (IE) refers to the variety of English that has developed in the Indian subcontinent. In India, it is spoken as an L2 by 129 million people (Census of India, 2011).  English is one of the two official languages, used in education, law, media, as a \emph{lingua franca} mainly for an educated elite in most metropolitan regions, and carries a high prestige value \citep{pandey201517, kachru1981english, tollefson2014language, kachru1983indianization}.
%citations 
Recent research comparing the segmental and suprasegmental properties of IE spoken in different parts of the country report many commonalities (see \cite{sirsa2013effects} for a review). This suggests that in spite of the varied L1s of its speakers, IE has a target phonology that is distinct from any of these, as well as from other native varieties of English. Thus, it is fruitful to think of regional variations as resulting from L1-influence on a common underlying target. Excepting a small minority of specific communities \citep{pandey201517, wells1982accents, coelho1997anglo}, there are very few L1 speakers of IE.\\

%%%%%%%%%%%%%%%%%%%%%%%do something%%%%%%%%%%%%%%%
Considering the language usage patterns suggested by this demography, the present study focuses on short-term phonetic interaction during mixed-language use, because:
\begin{enumerate}[label=(\roman*)]
	\item Since the population is multilingual, we expect long-term representations to be affected by multiple languages. 
	\item Given that examples of an IE phonology without L1 `influence' are very rare, it is more meaningful to think of cross-language transfer in L2 as relative to a speaker's own production in a given `baseline' condition.
\end{enumerate}

The socio-linguistic facts about English in India also suggest that for the present population, mixed-language processing of English and an L1 is an ecologically valid paradigm that is likely to be a part of everyday language experience. Thus, data from this population is valuable to the understanding of short-term phonetic interaction as it occurs in .

@@

The vowel system of IE contains the monopthongs \nt{I, i, E, e, \ae, @/2, a:, O, o, U, u}, represented by the lexical set KIT, FLEECE, DRESS, FACE, and TRAP, STRUT, PALM, LOT/CLOTH, GOAT, FOOT, GOOSE \citep{wells1982accents, masica1972sound}. %Their distribution in the F1XF2 space is shown in (insert table 1.7, Pandey). 
A single mid-central vowel corresponds to the categories \nt{2,@,3:}, which are treated as distinct in many native varieties of English \citep{nihalani1979indian,wells1982accents,hickey2005legacies,bansal1969intelligibility}. Since the English items used in this study are traditionally transcribed with \nt{2}, we use this symbol to indicate the mid-central vowel throughout.

@@

Existing research has distinguished changes in category \textsc{representations} due to the acquisition of multiple sound systems (cf. Speech Learning Model \citep{flege1995second,flege2007language}; Perceptual Assimilation Model-L2 \citep{best2007nonnative}), from interaction during accessing, processing, or articulation of these categories (cf. transfer vs interference \citep{grosjean2012attempt}; competence vs performance interference \citep{paradis1993linguistic}). 
Given their transient nature, dynamic changes in production during mixed-language use are generally attributed to the latter, e.g. online processing costs \citep{olson2013bilingual,tsui2019impact,vsimavckova2015immediate}, language mode \cite{simonet2014phonetic}, context-awareness \citep{khattab2013phonetic}. What triggers this interaction? \cite{olson2016role} argues that while cross-language phonetic effects are largely measured at the point of language switch, it could have two potential sources:
\begin{enumerate}
	\item The local point of switch itself
	\item Global co-activation of two languages -- bilingual language mode \citep{grosjean1998studying} 
\end{enumerate}

@@

\cite{grosjean1998studying} suggests that various aspects of the communicative setting, including exposure to (spoken or written) stimuli in multiple languages and awareness of the interlocutor's being bilingual, could trigger a bilingual mode. The participants in the present study are in an environment which largely contains mixed-language input, multilingual interlocutors, and no stable ambient language. Therefore we expect that phonetic transfer during everyday language use, if any, takes place in a bilingual mode.  To preserve ecological validity, we elicited both unilingual and switched utterances in a bilingual language mode. Any observed differences in this paradigm would arguably result from interaction during online processing.
